PANNEURO_analysis

### 2020-09-10 PANNEURO_analysis.Rmd
PANNEURO script and neuro chunks in analysis script

---
title: "PANCHANGE_analysis"
author: "Katie Thompson, Topher Huebel, Molly Davies, Kirstin Purves"
date: Sys.Date()
output:  
  html_document:
    df_print: paged
    toc: yes
    toc_depth: 2
    toc_float:
      collapsed: false
    number_sections: false
    highlight: monochrome
    theme: cerulean
code_folding: show

html_notebook:
  theme: cerulean
toc: yes
---

This is the analysis script to prepare the data to run analyses for the project "Anxiety, depression and trauma symptom change during the COVID-19 pandemic: retrospective versus objective assessment" - Young et al (2020)

Script written by K Purves, K Thompson, C Huebel and M Davies.
Email: kirstin.purves@kcl.ac.uk, katie.thompson@kcl.ac.uk, christopher.1.huebel@kcl.ac.uk, molly.davies@kcl.ac.uk

@@

```{r neurological frequencies}
dat.raw %>%
  filter(!is.na(startDate.coping)) %>%
  filter(!is.na(covid_test_positive)) %>%
  group_by(covid_test_positive) %>%
  select(
    starts_with("alsfrs"),
    -ends_with("numeric")
  ) %>% 
  colnames()
```

@@

```{r neurological frequencies}
dat.raw %>%
  filter(!is.na(startDate.coping)) %>%
  filter(!is.na(covid_test_positive)) %>%
  group_by(covid_test_positive) %>%
  select(
    starts_with("nmss"),
    -ends_with("numeric")
  ) %>% 
  colnames()
```

@@

```{r neurological frequencies}
dat.raw %>%
  filter(!is.na(startDate.coping)) %>%
  filter(!is.na(covid_test_positive)) %>%
  group_by(covid_test_positive) %>%
  select(
    starts_with("nmsw"),
    -ends_with("numeric")
  ) %>% 
  colnames()
```

@@

```{r neurological frequencies}
dat.raw %>%
  filter(!is.na(startDate.coping)) %>%
  filter(!is.na(covid_test_positive)) %>%
  group_by(covid_test_positive) %>%
  select(
    starts_with("tinnitus"),
    -ends_with("numeric")
  ) %>% 
  colnames()
```

@@

# Ethnicity_collapsed
```{r Ethnicity_collapsed per Sample prop per total}
Ethnicity_collapsed_per_Sample_prop_total <- dat %>%
  count(Ethnicity_collapsed, Sample) %>%
  mutate(Prop = round(n/sum(n), 2))

kable(Ethnicity_collapsed_per_Sample_prop_total)
Ethnicity_collapsed_per_Sample_prop_total
```

# Exclusions

Exclude GLAD/EDGI individuals that didn't take part in COPING
Exclude individuals with NA for the MHD categories
Individuals with NAs for age/gender have already been excluded in pre-processing script.

```{r exclusion criteria for diagnoses analysis}
dat <- dat.raw %>%
  filter(
    !is.na(depressive_disorders) &
      !is.na(eating_disorders) &
      !is.na(psychotic_disorders) &
      !is.na(obsessive_compulsive_disorders) &
      !is.na(autism_spectrum_disorder) &
      !is.na(mhd.mania_hypomania_bipolar_or_manicdepression_numeric) &
      !is.na(mhd.posttraumatic_stress_disorder_ptsd_numeric) &
      !is.na(control)
  ) %>% 
  drop_na(startDate.coping)

dim(dat.raw)
dim(dat)

```

# Collapsing categories

## age_category
```{r Inspect age_category}
dat %>%
  freq(age_category)

dat %>%
  freq(age_category_numeric)
```

```{r Collapse age_category}
dat <- dat %>%
  mutate(age_category_collapsed_numeric =
           case_when(
             age_category_numeric == 1 ~ 0,
             age_category_numeric == 2 ~ 1,
             age_category_numeric == 3 ~ 2,
             age_category_numeric == 4 ~ 3,
             age_category_numeric == 5 ~ 4,
             age_category_numeric == 6 ~ 5,
             age_category_numeric == 7 ~ 6,
             age_category_numeric == 8 ~ 7,
             age_category_numeric == 9 ~ 8,
             age_category_numeric == 10 ~ 8,
             age_category_numeric == 11 ~ 8,
             age_category_numeric == 12 ~ 8,
             age_category_numeric == 13 ~ 8

           )
  )

dat <- dat %>%
  mutate(age_category_collapsed =
           recode_factor(
             age_category_collapsed_numeric,
             `0` = "16-18",
             `1` = "19-25",
             `2` = "26-35",
             `3` = "36-45",
             `4` = "46-55",
             `5` = "56-65",
             `6` = "66-70",
             `7` = "71-75",
             `8` = "76+"
           )
  )

dat %>%
  freq(age_category_collapsed)
```

## Gender
```{r Inspect Gender}
dat %>%
  freq(Gender)

dat %>%
  freq(Gender_numeric)
```

```{r Collapse Gender}
dat <- dat %>%
  mutate(Gender_collapsed_numeric =
           case_when(
             Gender_numeric == 0 ~ 0,
             Gender_numeric == 1 ~ 1,
             Gender_numeric == 2 ~ 2,
             Gender_numeric == 3 ~ 2
           )
         )

dat <- dat %>%
  mutate(Gender_collapsed =
           recode_factor(
             Gender_collapsed_numeric,
             `0` = "Male",
             `1` = "Female",
             `2` = "Non-binary/Self-defined"
           )
         )

dat %>%
  freq(Gender_collapsed)
```

## Ethnicity
++MD: The ethnicity_numeric variable has more NAs (N = 14304) than the ethnicity categorical variable (N = 3568). Something strange must have happened during the conversion to numeric?
```{r Inspect Ethnicity}
dat %>%
  freq(Ethnicity)

dat %>%
  freq(Ethnicity_numeric)

```

++MD: Need to decide what to do with the Arab grouping.
++MD: Why is "Other ethnic group" recoded as Asian or Asian British?
```{r Collapse Ethnicity}
dat <- dat %>%
  mutate(Ethnicity_collapsed_numeric =
           case_when(
             Ethnicity_numeric == 1 ~ 0, #European
             Ethnicity_numeric == 2 ~ 1, #Mixed or multiple ethnic origins 
             Ethnicity_numeric == 3 ~ 2, #Asian or Asian British
             Ethnicity_numeric == 4 ~ 3, #African or African British
             Ethnicity_numeric == 5 ~ 3, #Arab - combined with African/African British
             Ethnicity_numeric == 6 ~ 4  #Other ethnic group
           )
  )

dat <- dat %>%
  mutate(Ethnicity_collapsed =
           recode_factor(
             Ethnicity_collapsed_numeric,
             `0` = "European",
             `1` = "Mixed or multiple ethnic origins",
             `2` = "Asian or Asian British",
             `3` = "African or African British",
             `4` = "Other"
           )
  )

dat %>%
  freq(Ethnicity_collapsed)
```

# Vectors for sum_score

```{r Vectors with sum_score_cols}
sum_score_cols <- dat %>%
  select(contains("sum_score")) %>%
  select(-ends_with("raw")) %>%
  colnames()

sum_score_cols

phq.sum_score_cols <- dat %>%
  select(contains("phq")) %>%
  select(contains("sum_score")) %>% 
  select(-ends_with("raw")) %>%
  select(-contains("8items")) %>%
  colnames()

phq.sum_score_cols

gad.sum_score_cols <- dat %>%
  select(contains("gad")) %>%
  select(contains("sum_score")) %>% 
  select(-ends_with("raw")) %>%
  colnames()

gad.sum_score_cols

ocir.sum_score_cols <- dat %>%
  select(contains("ocir")) %>%
  select(contains("sum_score")) %>% 
  select(-ends_with("raw")) %>%
  colnames()

ocir.sum_score_cols

pcl.sum_score_cols <- dat %>%
  select(contains("pcl")) %>%
  select(contains("sum_score")) %>% 
  select(-ends_with("raw")) %>%
  colnames()

pcl.sum_score_cols
```

# Vectors for diff_score

```{r Vectors with diff_score_cols}
diff_score_cols <- dat %>%
  select(contains("diff_score")) %>% 
  colnames()

diff_score_cols

phq.diff_score_cols <- dat %>%
  select(contains("phq")) %>%
  select(contains("diff_score")) %>% 
  colnames()

phq.diff_score_cols

gad.diff_score_cols <- dat %>%
  select(contains("gad")) %>%
  select(contains("diff_score")) %>% 
  colnames()

gad.diff_score_cols

ocir.diff_score_cols <- dat %>%
  select(contains("ocir")) %>%
  select(contains("diff_score")) %>% 
  colnames()

ocir.diff_score_cols
```

# ggplot theme
Set up ggplot theme for the plots
```{r ggplot theme}
theme_personal <-  theme(
    text = element_text(color = "black"),
    axis.title = element_text(color = "black"),
    axis.text = element_text(color = "black"),
    axis.title.y = element_blank(), 
    legend.background = element_blank(),
    legend.box.background = element_blank(),
    panel.background = element_blank(), 
    panel.grid.minor.x = element_blank(),
    panel.grid.major.x = element_line(
      colour = "gray",
      linetype = "dashed",
      size = 0.2
      ),
    axis.ticks = element_blank()
    )
```

# Psychiatric disorder hierarchy
- - - -
OLD VERSIONS:
control_numeric == 1 ~ 0,
depressive_disorders_numeric == 1 ~ 1,
anxiety_disorders_numeric == 1 ~ 2,
mhd.posttraumatic_stress_disorder_ptsd_numeric == 1 ~ 3,
eating_disorders_numeric == 1 ~ 4,
obsessive_compulsive_disorders_numeric == 1 ~ 5,
mhd.attention_deficit_hyperactivity_disorder_numeric == 1 ~ 6,
mhd.personality_disorder_numeric == 1 ~ 7,
autism_spectrum_disorder_numeric == 1 ~ 8,
mhd.mania_hypomania_bipolar_or_manicdepression_numeric == 1 ~ 9,
psychotic_disorders_numeric == 1 ~ 10

psychotic_disorders_numeric == 1 ~ 10
mhd.mania_hypomania_bipolar_or_manicdepression_numeric == 1 ~ 9,
autism_spectrum_disorder_numeric == 1 ~ 8,
mhd.personality_disorder_numeric == 1 ~ 7,
mhd.attention_deficit_hyperactivity_disorder_numeric == 1 ~ 6,
obsessive_compulsive_disorders_numeric == 1 ~ 5,
eating_disorders_numeric == 1 ~ 4,
mhd.posttraumatic_stress_disorder_ptsd_numeric == 1 ~ 3,
anxiety_disorders_numeric == 1 ~ 2,
depressive_disorders_numeric == 1 ~ 1,
control_numeric == 1 ~ 0,
- - - -

Disorder hierarchy
1: Psychotic disorders (create one with, create one without to test in analyses)
2: Bipolar disorder
3: MDD
4: PTSD
5: OCD
6: Eating disorders
7: Anxiety disorders

```{r Psychiatric disorders hierarchy including psychosis}
dat <- dat %>%
  mutate(
    Disorder_hierarchical_psychosis_numeric =
      case_when(
        psychotic_disorders_numeric == 1 ~ 7,
        mhd.mania_hypomania_bipolar_or_manicdepression_numeric == 1 ~ 6,
        depressive_disorders_numeric == 1 ~ 5,
        mhd.posttraumatic_stress_disorder_ptsd_numeric == 1 ~ 4,
        obsessive_compulsive_disorders_numeric == 1 ~ 3,
        eating_disorders_numeric == 1 ~ 2,
        anxiety_disorders_numeric == 1 ~ 1,
        control_numeric == 1 ~ 0
      )
  )

dat <- dat %>%
  mutate(
    Disorder_hierarchical_psychosis =
      recode_factor(
        Disorder_hierarchical_psychosis_numeric,
        `0` = "Controls",
        `1` = "Anxiety disorders",
        `2` = "Eating disorders",
        `3` = "OCDs",
        `4` = "PTSD",
        `5` = "Depressive disorders",
        `6` = "Bipolar disorders",
        `7` = "Psychotic disorders"
      )
  )

dat %>%
  freq(Disorder_hierarchical_psychosis)
```

```{r Psychiatric disorders hierarchy without psychosis}
dat <- dat %>%
  mutate(
    Disorder_hierarchical_numeric =
      case_when(
        mhd.mania_hypomania_bipolar_or_manicdepression_numeric == 1 ~ 6,
        depressive_disorders_numeric == 1 ~ 5,
        mhd.posttraumatic_stress_disorder_ptsd_numeric == 1 ~ 4,
        obsessive_compulsive_disorders_numeric == 1 ~ 3,
        eating_disorders_numeric == 1 ~ 2,
        anxiety_disorders_numeric == 1 ~ 1,
        control_numeric == 1 ~ 0
      )
  )

dat <- dat %>%
  mutate(
    Disorder_hierarchical =
      recode_factor(
        Disorder_hierarchical_numeric,
        `0` = "Controls",
        `1` = "Anxiety disorders",
        `2` = "Eating disorders",
        `3` = "OCDs",
        `4` = "PTSD",
        `5` = "Depressive disorders",
        `6` = "Bipolar disorders"
      )
  )

dat %>%
  freq(Disorder_hierarchical) 
dat %>% 
  freq(Disorder_hierarchical_psychosis)
```

Investigate NAs 
NAs are coded for individuals that don't have one of the listed disorders in the hierarchy, but who have another diagnosis that excludes them from being a control (e.g. personality disorder, ADHD, ASD). These individuals will be excluded from the analyses.
```{r NAs in diagnosis}
dat %>%
  filter(is.na(Disorder_hierarchical)) %>%
  freq(Sample)
```

Compare the GLAD/EDGI and COPING cohorts as a whole
+++CH: Diagnosis, ethnicity, gender, age

# age_category
```{r age_category per Sample prop per total}
age_category_per_Sample_prop_total <- dat %>%
  count(age_category, Sample) %>%
  mutate(Prop = round(n/sum(n), 2))

kable(age_category_per_Sample_prop_total)
age_category_per_Sample_prop_total
```

```{r age_category per Sample prop per Sample}
age_category_per_Sample_prop_Sample <- dat %>%
  group_by(Sample) %>%
  count(age_category, Sample) %>%
  mutate(Prop = round(n/sum(n), 2))

kable(age_category_per_Sample_prop_Sample)
age_category_per_Sample_prop_Sample
```

```{r age_category per Sample prop cross table}
age_category_per_Sample_prop_Sample_cross_table <- dat %>%
  group_by(Sample) %>%
  count(age_category, Sample) %>%
  mutate(Prop = round(n/sum(n), 2)) %>%
  pivot_wider(names_from = Sample, values_from = c(n, Prop))

kable(age_category_per_Sample_prop_Sample_cross_table)
age_category_per_Sample_prop_Sample_cross_table
```

```{r age_category per Sample count cross table}
age_category_per_Sample_count_cross_table <- dat %>%
  count(age_category, Sample) %>%
  spread(key = Sample, value = n) %>%
  mutate(
    EDGI = as.integer(replace_na(EDGI, 0)),
    GLAD = as.integer(replace_na(GLAD, 0)),
    NBR = as.integer(replace_na(NBR, 0)),
    RAMP = as.integer(replace_na(RAMP, 0))
  )

kable(age_category_per_Sample_count_cross_table)
age_category_per_Sample_count_cross_table
```

```{r age_category per Sample chisquare test}
age_category_per_Sample_chissquare_test <- dat %>%
  count(age_category, Sample) %>%
  spread(key = Sample, value = n) %>%
  select(-age_category) %>%
  mutate(
    EDGI = as.integer(replace_na(EDGI, 0)),
    GLAD = as.integer(replace_na(GLAD, 0)),
    NBR = as.integer(replace_na(NBR, 0)),
    RAMP = as.integer(replace_na(RAMP, 0))
  ) %>%
  chisq.test() %>%
  glance()

kable(age_category_per_Sample_chissquare_test)
age_category_per_Sample_chissquare_test
```

```{r age_category by Sample count plot}
age_category_by_Sample_count_plot <- age_category_per_Sample_prop_Sample %>%
  ggplot(mapping = aes(
    x = age_category,
    y = n)
  ) +
  geom_bar(
    aes(fill = Sample),
    stat="identity",
    binwidth = 1,
    position = position_dodge2(preserve = "single")
  ) +
  labs(
    x = "age_category",
    y = "Number of participants",
    fill = "Sample") +
  theme_personal +
  scale_fill_manual(
    values = COPINGpalette6
  ) +
  coord_flip()

age_category_by_Sample_count_plot
```

```{r age_category by Sample prop plot}
age_category_by_Sample_prop_plot <- age_category_per_Sample_prop_Sample %>%
  ggplot(mapping = aes(
    x = age_category,
    y = Prop)
  ) +
  geom_bar(
    aes(fill = Sample),
    stat="identity",
    binwidth = 1,
    position = position_dodge2(preserve = "single")
  ) +
  labs(
    x = "age_category",
    y = "Percentage per sample",
    fill = "Sample") +
  theme_personal +
  scale_fill_manual(
    values = COPINGpalette6
  ) +
  coord_flip()

age_category_by_Sample_prop_plot
```

age bar plot for whole sample (not grouped)

Create a nice summary dataset for age and display it as a bar plot

```{r age_category prop total (ungrouped)}

age_category_prop_total <- dat %>%
  count(age_category) %>%
  mutate(Prop = round(n/sum(n), 2))

# add a column to help order the factor
age_category_prop_total$order <- c(1,2,3,4,5,6,7,8,9,10,11,12,13)

# total responders
ageN <- sum(age_category_prop_total$n)

kable(age_category_prop_total)
age_category_prop_total

```

```{r age_category prop total (ungrouped) plot}
#Colour palette
RAMPagepalette <- c(
  "#00D9AC","#21D9B3","#4DD9BC","#78D9C5","#8ED9C9","#A3D9CE",
  "#CFD9D7",
  "#DAD0BA","#DAC38F","#DAC38F","#DAB663" ,"#DAA937","#DA9C0C")

#Bar plot
plot.age <-

  age_category_prop_total %>%
    mutate(age_category = fct_reorder(age_category,order)) %>%

  ggplot(aes(x = age_category)) +
  geom_bar(aes(y = n, color=age_category, fill=age_category, alpha = 1),
           stat ="summary") +
  scale_fill_manual(values =  RAMPagepalette) +
  scale_color_manual(values =  RAMPagepalette) +
  labs(x = "",
       y = "Total count",
       title = paste("Number of participants in each age category",sep=" "),
       subtitle = paste("n = ", ageN,
                        sep = ""),
       fill = "Age category") +
   guides(alpha = FALSE, color = FALSE, fill = FALSE) +
  scale_x_discrete(limits = (levels(age_category_prop_total$age_category))) +
  theme_classic2() +
   theme(axis.text.x = element_text(colour="black", size = 12,angle=45,
                                    vjust=0.5),
        axis.text.y = element_text(colour="black", size = 12))

plot.age

```

# Ethnicity_collapsed
```{r Ethnicity_collapsed per Sample prop per total}
Ethnicity_collapsed_per_Sample_prop_total <- dat %>%
  count(Ethnicity_collapsed, Sample) %>%
  mutate(Prop = round(n/sum(n), 2))

kable(Ethnicity_collapsed_per_Sample_prop_total)
Ethnicity_collapsed_per_Sample_prop_total
```

```{r Ethnicity_collapsed per Sample prop per Sample}
Ethnicity_collapsed_per_Sample_prop_Sample <- dat %>%
  group_by(Sample) %>%
  count(Ethnicity_collapsed, Sample) %>%
  mutate(Prop = round(n/sum(n), 2))

kable(Ethnicity_collapsed_per_Sample_prop_Sample)
Ethnicity_collapsed_per_Sample_prop_Sample
```

```{r Ethnicity_collapsed per Sample prop cross table}
Ethnicity_collapsed_per_Sample_prop_Sample_cross_table <- dat %>%
  group_by(Sample) %>%
  count(Ethnicity_collapsed, Sample) %>%
  mutate(Prop = round(n/sum(n), 2)) %>%
  pivot_wider(names_from = Sample, values_from = c(n, Prop))

kable(Ethnicity_collapsed_per_Sample_prop_Sample_cross_table)
Ethnicity_collapsed_per_Sample_prop_Sample_cross_table
```

```{r Ethnicity_collapsed per Sample count cross table}
Ethnicity_collapsed_per_Sample_count_cross_table <- dat %>%
  count(Ethnicity_collapsed, Sample) %>%
  spread(key = Sample, value = n) %>%
  mutate(
    EDGI = as.integer(replace_na(EDGI, 0)),
    GLAD = as.integer(replace_na(GLAD, 0)),
    NBR = as.integer(replace_na(NBR, 0)),
    RAMP = as.integer(replace_na(RAMP, 0))
  )

kable(Ethnicity_collapsed_per_Sample_count_cross_table)
Ethnicity_collapsed_per_Sample_count_cross_table
```

```{r Ethnicity_collapsed per Sample chisquare test}
Ethnicity_collapsed_per_Sample_chissquare_test <- dat %>%
  count(Ethnicity_collapsed, Sample) %>%
  spread(key = Sample, value = n) %>%
  select(-Ethnicity_collapsed) %>%
  mutate(
    EDGI = as.integer(replace_na(EDGI, 0)),
    GLAD = as.integer(replace_na(GLAD, 0)),
    NBR = as.integer(replace_na(NBR, 0)),
    RAMP = as.integer(replace_na(RAMP, 0))
  ) %>%
  chisq.test() %>%
  glance()

kable(Ethnicity_collapsed_per_Sample_chissquare_test)
Ethnicity_collapsed_per_Sample_chissquare_test
```

```{r Ethnicity_collapsed by Sample count plot}
Ethnicity_collapsed_by_Sample_count_plot <- Ethnicity_collapsed_per_Sample_prop_Sample %>%
  ggplot(mapping = aes(
    x = Ethnicity_collapsed,
    y = n)
  ) +
  geom_bar(
    aes(fill = Sample),
    stat="identity",
    binwidth = 1,
    position = position_dodge2(preserve = "single")
  ) +
  labs(
    x = "Ethnicity_collapsed",
    y = "Number of participants",
    fill = "Sample") +
  theme_personal +
  scale_fill_manual(
    values = COPINGpalette6
  ) +
  coord_flip()

Ethnicity_collapsed_by_Sample_count_plot
```

```{r Ethnicity_collapsed by Sample prop plot}
Ethnicity_collapsed_by_Sample_prop_plot <- Ethnicity_collapsed_per_Sample_prop_Sample %>%
  ggplot(mapping = aes(
    x = Ethnicity_collapsed,
    y = Prop)
  ) +
  geom_bar(
    aes(fill = Sample),
    stat="identity",
    binwidth = 1,
    position = position_dodge2(preserve = "single")
  ) +
  labs(
    x = "Ethnicity_collapsed",
    y = "Percentage per sample",
    fill = "Sample") +
  theme_personal +
  scale_fill_manual(
    values = COPINGpalette6
  ) +
  coord_flip()

Ethnicity_collapsed_by_Sample_prop_plot
```

### pie charts

```{r Ethnicity_collapsed prop per total summary}

Ethnicity_collapsed_prop_Sample <- dat %>%
  count(Ethnicity_collapsed) %>%
  drop_na() %>%
  mutate(freq = round(n/sum(n), 2))

# total responders
ethN <- sum(Ethnicity_collapsed_prop_Sample$n)

kable(Ethnicity_collapsed_prop_Sample)
Ethnicity_collapsed_prop_Sample

```

pie chart with only European label

```{r ethnicity pie chart biggest group label}

eth_pie_alllabel <-

  Ethnicity_collapsed_prop_Sample %>%
   mutate(Ethnicity_collapsed = fct_reorder(Ethnicity_collapsed,-n)) %>%

  ggplot(aes(x="", y=freq, fill=Ethnicity_collapsed)) +
  geom_bar(width = 1, stat = "identity") +
  scale_fill_manual(values =  COPINGpalette6) +
  scale_color_manual(values =  COPINGpalette6) +
  coord_polar("y",start=0) +
  theme_void() +
   labs(title = paste("RAMP/COPING sample self reported ethnicity",
                      sep=" "),
        subtitle = paste("n =", sum(Ethnicity_collapsed_prop_Sample$n),
                         sep = " "),
        fill = "Self reported ethnicity")  +
    geom_text(aes(y = c(0.5),
                  label= c(paste0(round((freq[1]*100),0),"%"))), #Frequency of European only
             size=5,
             face="bold"
             )

eth_pie_alllabel

```

pie chart - no labels

```{r ethnicity pie chart no label}

eth_pie <-

    Ethnicity_collapsed_prop_Sample %>%
   mutate(Ethnicity_collapsed = fct_reorder(Ethnicity_collapsed,-n)) %>%

  ggplot(aes(x="", y=freq, fill=Ethnicity_collapsed)) +
  geom_bar(width = 1, stat = "identity") +
  scale_fill_manual(values =  COPINGpalette6) +
  scale_color_manual(values =  COPINGpalette6) +
  coord_polar("y",start=0) +
  theme_void() +
   labs(title = paste("RAMP/COPING sample self reported ethnicity",
                      sep=" "),
        subtitle = paste("n =", sum(Ethnicity_collapsed_prop_Sample$n),
                         sep = " "),
        fill = "Self reported ethnicity")

eth_pie

```

# Gender_collapsed
```{r Gender_collapsed per Sample prop per total}
Gender_collapsed_per_Sample_prop_total <- dat %>%
  count(Gender_collapsed, Sample) %>%
  mutate(Prop = round(n/sum(n), 2))

kable(Gender_collapsed_per_Sample_prop_total)
Gender_collapsed_per_Sample_prop_total
```

```{r Gender_collapsed per Sample prop per Sample}
Gender_collapsed_per_Sample_prop_Sample <- dat %>%
  group_by(Sample) %>%
  count(Gender_collapsed, Sample) %>%
  mutate(Prop = round(n/sum(n), 2))

kable(Gender_collapsed_per_Sample_prop_Sample)
Gender_collapsed_per_Sample_prop_Sample
```

```{r Gender_collapsed per Sample prop cross table}
Gender_collapsed_per_Sample_prop_Sample_cross_table <- dat %>%
  group_by(Sample) %>%
  count(Gender_collapsed, Sample) %>%
  mutate(Prop = round(n/sum(n), 2)) %>%
  pivot_wider(names_from = Sample, values_from = c(n, Prop))

kable(Gender_collapsed_per_Sample_prop_Sample_cross_table)
Gender_collapsed_per_Sample_prop_Sample_cross_table
```

```{r Gender_collapsed per Sample count cross table}
Gender_collapsed_per_Sample_count_cross_table <- dat %>%
  count(Gender_collapsed, Sample) %>%
  spread(key = Sample, value = n) %>%
  mutate(
    EDGI = as.integer(replace_na(EDGI, 0)),
    GLAD = as.integer(replace_na(GLAD, 0)),
    NBR = as.integer(replace_na(NBR, 0)),
    RAMP = as.integer(replace_na(RAMP, 0))
  )

kable(Gender_collapsed_per_Sample_count_cross_table)
Gender_collapsed_per_Sample_count_cross_table
```

```{r Gender_collapsed per Sample chisquare test}
Gender_collapsed_per_Sample_chissquare_test <- dat %>%
  count(Gender_collapsed, Sample) %>%
  spread(key = Sample, value = n) %>%
  select(-Gender_collapsed) %>%
  mutate(
    EDGI = as.integer(replace_na(EDGI, 0)),
    GLAD = as.integer(replace_na(GLAD, 0)),
    NBR = as.integer(replace_na(NBR, 0)),
    RAMP = as.integer(replace_na(RAMP, 0))
  ) %>%
  chisq.test() %>%
  glance()

kable(Gender_collapsed_per_Sample_chissquare_test)
Gender_collapsed_per_Sample_chissquare_test
```

```{r Gender_collapsed by Sample count plot}
Gender_collapsed_by_Sample_count_plot <- Gender_collapsed_per_Sample_prop_Sample %>%
  ggplot(mapping = aes(
    x = Gender_collapsed,
    y = n)
  ) +
  geom_bar(
    aes(fill = Sample),
    stat="identity",
    binwidth = 1,
    position = position_dodge2(preserve = "single")
  ) +
  labs(
    x = "Gender_collapsed",
    y = "Number of participants",
    fill = "Sample") +
  theme_personal +
  scale_fill_manual(
    values = COPINGpalette6
  ) +
  coord_flip()

Gender_collapsed_by_Sample_count_plot
```

```{r Gender_collapsed by Sample prop plot}
Gender_collapsed_by_Sample_prop_plot <- Gender_collapsed_per_Sample_prop_Sample %>%
  ggplot(mapping = aes(
    x = Gender_collapsed,
    y = Prop)
  ) +
  geom_bar(
    aes(fill = Sample),
    stat="identity",
    binwidth = 1,
    position = position_dodge2(preserve = "single")
  ) +
  labs(
    x = "Gender_collapsed",
    y = "Percentage per sample",
    fill = "Sample") +
  theme_personal +
  scale_fill_manual(
    values = COPINGpalette6
  ) +
  coord_flip()

Gender_collapsed_by_Sample_prop_plot
```

### pie charts

```{r Gender_collapsed prop per total}

Gender_collapsed_prop_total <- dat %>%
  count(Gender_collapsed) %>%
  mutate(freq = round(n/sum(n), 2))

kable(Gender_collapsed_prop_total)
Gender_collapsed_prop_total

```

```{r gender pie chart - all labels}

gen_pie_alllabel <- 
  ggplot(Gender_collapsed_prop_total, aes(x="", y=freq, fill=Gender_collapsed)) +
  geom_bar(width = 1, stat = "identity") +
  scale_fill_manual(values =  COPINGpalette5) +
  scale_color_manual(values =  COPINGpalette5) +
  coord_polar("y",start=0) +
  theme_void() +
   labs(title = paste("RAMP/COPING sample self reported gender",
                      sep=" "),
        subtitle = paste("n", sum(Gender_collapsed_prop_total$n),
                         sep = " "),
        fill = "Self reported gender")  +
    geom_text(aes(y = c(0.85,0.35,0),
                  x = c(1.2,1.2,1.8),
                  label= c(paste0(round((freq[1:2]*100),0),"%"),paste0(round((freq[3]*100),1),"%"))),
             size=5,
             face="bold"
             )

gen_pie_alllabel

```

pie chart - only labelling two largest groups (male and female)

```{r gender pie chart - male/female labels only}

gen_pie <- 
  ggplot(Gender_collapsed_prop_total, aes(x="", y=freq, fill=Gender_collapsed)) +
  geom_bar(width = 1, stat = "identity") +
  scale_fill_manual(values =  COPINGpalette5) +
  scale_color_manual(values =  COPINGpalette5) +
  coord_polar("y",start=0) +
  theme_void() +
   labs(title = paste("RAMP/COPING sample self reported gender",
                      sep=" "),
        subtitle = paste("n", sum(Gender_collapsed_prop_total$n),
                         sep = " "),
        fill = "Self reported gender")  +
    geom_text(aes(y = c(0.85,0.35,0),
                  x = c(1.2,1.2,1.8),
                  label= c(paste0(round((freq[1:2]*100),0),"%"),"")),
             size=5,
             face="bold"
             )

gen_pie

```

pie chart -  no labels

```{r gender pie chart - no labels }

gen_pie_nolabel <- 
  ggplot(Gender_collapsed_prop_total, aes(x="", y=freq, fill=Gender_collapsed)) +
  geom_bar(width = 1, stat = "identity") +
  scale_fill_manual(values =  COPINGpalette5) +
  scale_color_manual(values =  COPINGpalette5) +
  coord_polar("y",start=0) +
  theme_void() +
   labs(title = paste("RAMP/COPING sample self reported gender",
                      sep=" "),
        subtitle = paste("n", sum(Gender_collapsed_prop_total$n),
                         sep = " "),
        fill = "Self reported gender")

gen_pie_nolabel

```

# highest_education
```{r highest_education per Sample prop per total}
highest_education_per_Sample_prop_total <- dat %>%
  count(highest_education, Sample) %>%
  mutate(Prop = round(n/sum(n), 2))

kable(highest_education_per_Sample_prop_total)
highest_education_per_Sample_prop_total
```

```{r highest_education per Sample prop per Sample}
highest_education_per_Sample_prop_Sample <- dat %>%
  group_by(Sample) %>%
  count(highest_education, Sample) %>%
  mutate(Prop = round(n/sum(n), 2))

kable(highest_education_per_Sample_prop_Sample)
highest_education_per_Sample_prop_Sample
```

```{r highest_education per Sample prop cross table}
highest_education_per_Sample_prop_Sample_cross_table <- dat %>%
  group_by(Sample) %>%
  count(highest_education, Sample) %>%
  mutate(Prop = round(n/sum(n), 2)) %>%
  pivot_wider(names_from = Sample, values_from = c(n, Prop))

kable(highest_education_per_Sample_prop_Sample_cross_table)
highest_education_per_Sample_prop_Sample_cross_table
```

```{r highest_education per Sample count cross table}
highest_education_per_Sample_count_cross_table <- dat %>%
  count(highest_education, Sample) %>%
  spread(key = Sample, value = n) %>%
  mutate(
    EDGI = as.integer(replace_na(EDGI, 0)),
    GLAD = as.integer(replace_na(GLAD, 0)),
    NBR = as.integer(replace_na(NBR, 0)),
    RAMP = as.integer(replace_na(RAMP, 0))
  )

kable(highest_education_per_Sample_count_cross_table)
highest_education_per_Sample_count_cross_table
```

```{r highest_education per Sample chisquare test}
highest_education_per_Sample_chissquare_test <- dat %>%
  count(highest_education, Sample) %>%
  spread(key = Sample, value = n) %>%
  select(-highest_education) %>%
  mutate(
    EDGI = as.integer(replace_na(EDGI, 0)),
    GLAD = as.integer(replace_na(GLAD, 0)),
    NBR = as.integer(replace_na(NBR, 0)),
    RAMP = as.integer(replace_na(RAMP, 0))
  ) %>%
  chisq.test() %>%
  glance()

kable(highest_education_per_Sample_chissquare_test)
highest_education_per_Sample_chissquare_test
```

>>>>

---
title: "PANNEURO_analysis"
author: "Christopher Huebel"
date: "Sys.Date()"
output:
  word_document:
    toc: yes
    toc_depth: '2'
  pdf_document:
    toc: yes
    toc_depth: '2'
  html_document:
    df_print: paged
    highlight: monochrome
    number_sections: no
    theme: cerulean
    toc: yes
    toc_depth: 2
    toc_float:
      collapsed: no
html_notebook:
  theme: cerulean
code_folding: show
toc: yes
---

Frequencies of neurological symptoms in participants of the COPING/RAMP studies

Filtered by participants that self-reported a swap or antibody test

and that answered the neurological questionnaire

Script written by C Huebel
Email: christopher.1.huebel@kcl.ac.uk

@@

```{r neurological frequencies alsfrs}
dat.raw %>%
  filter(!is.na(startDate.coping)) %>%
  filter(!is.na(covid_test_positive)) %>%
  group_by(covid_test_positive) %>%
  select(
    starts_with("alsfrs"),
    -ends_with("numeric")
  ) %>% 
  colnames()
```

@@

```{r neurological frequencies nmss}
dat.raw %>%
  filter(!is.na(startDate.coping)) %>%
  filter(!is.na(covid_test_positive)) %>%
  group_by(covid_test_positive) %>%
  select(
    starts_with("nmss"),
    -ends_with("numeric")
  ) %>% 
  colnames()
```

@@

```{r neurological frequencies nmsw}
dat.raw %>%
  filter(!is.na(startDate.coping)) %>%
  filter(!is.na(covid_test_positive)) %>%
  group_by(covid_test_positive) %>%
  select(
    starts_with("nmsw"),
    -ends_with("numeric")
  ) %>% 
  colnames()
```

@@

```{r neurological frequencies tinnitus}
dat.raw %>%
  filter(!is.na(startDate.coping)) %>%
  filter(!is.na(covid_test_positive)) %>%
  group_by(covid_test_positive) %>%
  select(
    starts_with("tinnitus"),
    -ends_with("numeric")
  ) %>% 
  colnames()
```

@@
